---
title: "Assignment 2.4"
subtitle: "CPSC 292 Section -- (Instructor last name)"
author: "Your Name"
date: "00/00/2022"
output: html_notebook
---


## Assignment 2.4

```{r, include=TRUE}
# Loading required libraries

```

### Question 1

*Download the ElephantsMF.csv data file associated with the assignment. This data set is from Kaggle user Mostafa Elseidy \href{https://www.kaggle.com/mostafaelseidy/elephantsmf}{https://www.kaggle.com/mostafaelseidy/elephantsmf}.  According to the details, this data set represents data on 288 African elephants that are greater than 2 years old and are a part of the publication referenced below.  Load this data set into your R Markdown file.  *


```{r, include=TRUE}

```

### Question 2

*Answer the following using a line of code and a written explanation.*

*a. What are the columns in this data set and what data class are each?*
```{r, include=TRUE}

```


*b. What is the mean and standard deviations of the heights of elephants in the data set?*

```{r, include=TRUE}

```



*c. What is age range of the elephants in the data set? Are there more young or old elephants in the data set, or is the representation pretty even? (Hint: check the quartiles and use stripchart() to look at the distribution.)*

```{r, include=TRUE}

```


### Quesion 3: 

*Is there a height difference between male and female elephants in this data set? How can you tell?*

```{r, include=TRUE}

```

### Question 4: 

*Is there a relationship between height and age of these elephants? How can you tell?*


```{r, include=TRUE}

```
