---
title: "Skill Check 1"
date: "Fall 2022"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

This Skill Check is an individual assessment and you should not receive or offer help on it from any other human.

You may use any resource, either online or physical, to complete the work. This includes:

 - Any help forum or website (e.g. StackOverflow) questions that already exist. (You MAY NOT ask a question on a forum and then receive specific help from a person.)
 - Any notes, code, slides, papers, or previous feedback from the instructor.
 - Any books, online or physical.
 - Scholarly works such as papers.
 
You may NOT use:

 - Help from any other student or person. This is an individual assessment.
 - Asking for help on specific questions. (It is fine to ask for clarification!)
 - Help from homework websites such as Course Hero or Chegg.
 
You may choose to answer any one part of each question. You must get at least one correct for each question to get full credit. 

### Question 1

Pick one of the following questions to answer in the box below. 

 A. In the command line, how could you take the square root of the number 128? Write the code.
 
 B. In the command line, how could you assign the value 3 to the variable name bloop? Write the code.
 
 C. In the command line, how would you divide the number 3 into 149? Write the code.
 
```{r}


```

### Question 2

In the command line, how would you find the help documentation on the function mean()? Write the code here.

```{r}

```


## Question 3

Explain the syntax of an R function. Describe each part and what role it plays.

*Your answer here*

### Question 4

Pick two of the following questions and answer in the box below.

 A. In the command line, how would you use the `assign()` function to assign the value 8 to the variable y? Write the code. 
 
 B. In the command line, how could you assign the numbers 8.2 and 4 to a variable x? Write the code.
 
 C. In the command line, how would you create a vector that has length 15 in which all positions have value 1? Write the code.
 
 D. In the command line, how would you create a vector that is a sequence of numbers 4 through 13? Write the code.

```{r}

```

#### Question 5

Pick one of the following questions and answer it in the box below.

 A. Explain the tasks that all computers must accomplish. 
 
 B. Explain the difference between memory and storage.
 
 C. Write out the number 41 in binary. 
 
*Your answer here*






