---
title: "Lecture 2.4 Notebook"
author: Lindsay Waldrop
date: 10/3/2022
output: html_notebook
---

```{r,include=TRUE}
# Loading required libraries
# Be sure to install these before running this file!!

```

## Scripting & Troubleshooting Compound Code

#### Example: Blue Crab Graphing

In my dissertation, I made models of crab antennae hair arrays and measured fluid speed through spaces in between the hairs. I had a bunch of conditions: splayed or clumped arrays, backward or forward facing, and how fast external fluid was moving. 

I need to visualize these data and decide which factors are important. The data are in `bluecrab.csv` in the data zip.

```{r}
# Loading and cleaning data
crab.dat <- read.csv("data/bluecrab.csv")
crab.dat$Orient <- factor(crab.dat$Orient)
crab.dat$Model <- factor(crab.dat$Model)
#crab.dat$current <- factor(crab.dat$current)
```

Graphs: 

 Fluid speed versus current.
```{r}
plot(crab.dat$current, crab.dat$Speed, 
     pch=as.numeric(crab.dat$Orient),
     col=as.numeric(crab.dat$Orient),
     xlab = "External current (m/s)", 
     ylab = "Fluid speed in array (m/s)")
legend("topleft", legend = c("Backwards", "Forwards"), pch = c(1, 2), col = c(1, 2))
```

Speed versus hair splay:
```{r}
plot(crab.dat$Model, crab.dat$Speed, 
     xlab = "Hair splay model", 
     ylab = "Fluid speed in arary (m/s)")

```



#### Check Your Understanding

Import the data set in `tomatohaul2021.csv` as tomato.data. Graph tomato harvest (in lbs) versus date. Make the paste tomatoes blue and the heirloom tomatoes red. Be sure there are both points and lines on the graph. Include a legend. Add an additional element! 

```{r}

```

