---
title: "Assignment 3.3"
subtitle: "CPSC 292 Section -- (Instructor last name)"
author: "Your Name"
date: "10/28/2022"
output: html_notebook
---

## Assignment 3.3

```{r, include=TRUE}
# Loading required libraries
 

```

### Question 1

*Write a function that returns the correct sound that a user-specified animal makes. Include at
least four animals. Do not use switch() for this question.*

```{r}


```


### Question 2

*Write a function that acts like a magic 8 ball. It should take a question as an argument and return an answer (randomly) as output. Include a sample question that calls the function to demonstrate.*

```{r, include=TRUE}
magic8 <- function(question){
  print(question)
  a <-as.character(sample(20,1))
  answer<-switch(a,
         "1" = "It is certain.",
         "2" = "It is decidedly so.",
         "3" = "Without a doubt.",
         "4" = "Yes definitely.",
         "5" = "You may rely on it.",
         "6" = "As I see it, yes.",
         "7" = "Most likely.",
         "8" = "Outlook good.",
         "9" = "Yes.",
         "10" = "Signs point to yes.",
         "11" = "Reply hazy, try again.",
         "12" = "Ask again later.",
         "13" = "Better not tell you now.",
         "14" = "Cannot predict now.",
         "15" = "Concentrate and ask again.",
         "16" = "Don't count on it.",
         "17" = "My reply is no.",
         "18" = "My sources say no.",
         "19" = "Outlook not so good.",
         "20" = "Very doubtful.",
         "Oops, try again."
         )
  print(answer)
}

magic8("Sample Question?")

```

I got the responses from the Wikipedia page for the Magic 8 Ball: https://en.wikipedia.org/wiki/Magic_8-Ball 