---
title: "Assignment 3.2"
subtitle: "CPSC 292 Section -- (Instructor last name)"
author: "Your Name"
date: "10/28/2022"
output: html_notebook
---

## Assignment 3.2 

```{r, include=TRUE}
# Loading required libraries

```

### Question 1

*Complete the following steps:*

 a. *Generate 10 random, whole numbers using `sample()` and store them in a vector named `hop`.*

```{r, include=TRUE}

```
 
 b.  *Create a function that will take a single number \verb+x+ and return the results of the following calculation: $2x^2 -3x +12$*

```{r, include=TRUE}

```

 c. *Use your function to calculate the result of each element of `hop` individually. Store the result in a new vector called `hop.up1`.  Note: you must do this to each of the ten elements individually. Hint: create `hop.up1` first as a vector of `NA`'s. *
 
```{r, include=TRUE}

```

 d. *Now use your function to perform a vectorized calculation of the equation on `hop` and store this as `hop.up2`.*
 
```{r, include=TRUE}

```

 e.  *Last, use `all.equal()` to test whether or not `hop.up1` and `hop.up2` are the same. (They should be!)*
 
```{r, include=TRUE}

```

### Question 2

*Make a function that reports the data in `state.x77` (in `data("state")`) for a specific state defined by the user.*

```{r, include=TRUE}

```