---
title: "Lecture 2.8 Notebook"
author: Lindsay Waldrop
date: 10/17/2022
output: html_notebook
---

```{r setup}
# Loading required libraries
# Be sure to install these before running this file!!
library(ggplot2)
```

## Best Practices in Data Visualization

#### Check Your Understanding

```{r}
# Load the data set
data("diamonds")
force(diamonds)
```

About the diamonds data set (part of the ggplot2 library): https://ggplot2.tidyverse.org/reference/diamonds.html

Are flawless diamonds on average smaller than those with inclusions? 

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) + 
  geom_point(shape = 19) + 
  xlab("Carat")  + ylab("Price ($)") +
  theme_minimal()
```

How can we improve this visualization?

Remember: 

 1. _Attention is Queen._ What is the most important point to draw your viewer’s attention to? How can you create a relationship that shows this point?
 2. _Reduce demand on working memory._ Is there anything not relevant in your graph that you could exclude? Are your graph elements easy to understand and visually process?
 
Put improved graphic here: 
```{r}
# your improved graphic
```
