---
title: "Continous x and y data"
output: html_notebook
---
## geom_point
It is used as a means to change the graph or object being visualized. It adds a layer of points to the plots to create a scatterplot. And it is part of the ggplot2 which has many more geom functions and geom_point is one of them.

```{r}
data(mtcars)
p <-ggplot(mtcars, aes(x= mpg, y= hp))
p + geom_point(size = 4)
p + geom_point(shape= 8)
p + geom_point(color= "purple")
p + geom_point(pch = 19, color = "light green", fill = "black")
```
## geom_line
Geom_line() allows for modification of the observed data in the x-axis

```{r}
library(ggplot2)
data("Loblolly")
ggplot(Loblolly, aes(x=height, y=age, color=factor(Seed))) +
  ggtitle("Height vs Age") +
  geom_point()+
  geom_line(color="blue")
```


```{r}
library(ggplot2)
data("Loblolly")
ggplot(Loblolly, aes(x=height, y=age, color=factor(Seed))) +
  ggtitle("Height vs Age") +
  geom_point()+
  geom_line(color="blue", linetype=2)
```


```{r}
library(ggplot2)
data("Loblolly")
ggplot(Loblolly, aes(x=height, y=age, color=factor(Seed))) +
  ggtitle("Height vs Age") +
  geom_point() +
  geom_line(color="darkblue", linetype=3, size=2)
```

## geom_qq_line

A normal geom_qq plot shows the distribution of data against the expected natural distribution.It is used when comparing it to a bell shaped curve. The closer it is to the line, the more normally distributed data it is.

```{r}
library(ggplot2)

data("Loblolly")

ggplot(Loblolly, aes(sample= height, color = factor(Seed))) +
  xlab ("Theoretical") +
  ylab ("Height") +
  ggtitle ("Theoretical vs. Actual Height of Trees") + 
  stat_qq() +
  stat_qq_line()
```

