---
title: "Patchwork"
output: html_notebook
date: 10/10/22
---

The goal of patchwork is to make it simple to combine separate ggplots into the same graphic complex layouts. 
What we need to get started:
In order to successfully join 2 or more plots, the packages ggplot2 and patchwork need to be installed.

```{r}
library('patchwork') 
library('ggplot2')
```

Example #1
```{r}
data("iris")
```
First we start off by deciding our x and y varaibles. 
In p1 we useSepal.Length, Sepal.Width #In p2 and p3 we use Species, Sepal.Width 
```{r}
p1 <- ggplot(iris) + geom_point(aes(Sepal.Length, Sepal.Width, col= Species)) 

p1

p2 <- ggplot(iris, aes(x = Species, y = Sepal.Width,fill = Species)) +geom_bar(stat = "identity") 
p2 

p3 <- ggplot(iris) + geom_boxplot(aes(Species, Sepal.Width, col= Species)) 

p3
```
Now that we have all our different plots we can see how we can use patchwork to view multiple at a time The simplest way to use patchwork is by adding plots together
```{r}
p1+p2 
```
Plots can be stacked on top of each other using "/" and "|" is used to place the plots next to each other. 
```{r}
p2+p3/p1 
```
The next one shows how we can print the plots to lay on top of one another 
```{r}
ptop <- p1 +                  
      inset_element(p2, 0.01, 0.01, 0.7, 0.5) +
      inset_element(p3, 0.4, 0.6, 0.99, 0.99)

ptop
```
Often plots need to have titles, captions, tags, etc. This can be added to the plots by adding the function plot_annotation() to the original patchwork function. 
```{r}
title <- p1+p2/p3 

title + plot_annotation( title = 'The truth about the flower iris', subtitle = 'These 3 plots will reveal how widths and lengths of sepal vary through species',caption = 'Disclaimer')

title + plot_annotation(tag_levels = '1')
```
These tools can help us view a variety of data at once and help ourselves organize each graph through simple coding. When wanting to see multiple graphs at once, patchwork is the best tool to use.

What type of data is used: Since patchwork joins 2 or more plots, it uses continuous, categorical, and discrete data.


References:
Pendersen, Thomas L., "patchwork:The Composer of Plots." Github, August 2022, https://github.com/thomasp85/patchwork#patchwork. 

https://cran.r-project.org/web/packages/patchwork/vignettes/patchwork.html

https://statisticsglobe.com/patchwork-r-package


